---
title: "p8105_hw6_zc2691"
author: "Zhaohua Chunyu"
date: "2022-11-29"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1
### Problem 1

To obtain a distribution for $\hat{r}^2$, we'll follow basically the same procedure we used for regression coefficients: draw bootstrap samples; the a model to each; extract the value I'm concerned with; and summarize. Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r weather_df, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

In this example, the $\hat{r}^2$ value is high, and the upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for $R^2$, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, because the shape isn't symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.

We can produce a distribution for $\log(\beta_0 * \beta1)$ using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

As with $r^2$, this distribution is somewhat skewed and has some outliers. 

The point of this is not to say you should always use the bootstrap -- it's possible to establish "large sample" distributions for strange parameters / values / summaries in a lot of cases, and those are great to have. But it is helpful to know that there's a way to do inference even in tough cases. 


## Problem 2
```{r}
homicide_df = read_csv("data/homicide-data.csv")
```

Create a `city_state` variable and a binary variable `is_resolved` indicating whether the homicide is solved.

```{r}
homicide_df = 
  homicide_df %>% 
  mutate(city_state = str_c(city, "_", state)) %>% 
  filter(
    !city_state %in% c("Dallas_TX", "Phoenix_AZ", "Kansas City_MO", "Tulsa_AL"),
    victim_race %in% c("White", "Black")
  ) %>% 
  mutate (is_resolved = disposition == "Closed by arrest",
          victim_age = as.numeric(victim_age))
```

For the city of `Baltimore_MD`, use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_fit = 
  homicide_df %>% 
  filter(city_state == "Baltimore_MD") %>%
  mutate(is_resolved = as.numeric(is_resolved)) %>% 
  glm(is_resolved ~ victim_age + victim_sex + victim_race, data = ., family = "binomial") %>% 
  broom::tidy()  

baltimore_fit_log_result = 
  baltimore_fit %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(OR_estimate = exp(estimate),
         OR_ci_lower = exp(estimate + qnorm(0.025) * std.error),
         OR_ci_upper = exp(estimate + qnorm(0.975)* std.error)) %>% 
  select(term, OR_estimate, OR_ci_lower, OR_ci_upper)  

baltimore_fit_log_result %>% knitr::kable(digits = 3)
```

Fit `glm` for each of the cities and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. 

```{r}
all_city_nested = 
  homicide_df %>% 
  mutate(is_resolved = as.numeric(is_resolved)) %>% 
  select(city_state, is_resolved, victim_age, victim_sex, victim_race) %>% 
  nest(data = is_resolved:victim_race)

fit_log_result = 
  all_city_nested %>% 
  mutate(
    fit = purrr::map(.x = data, ~glm(is_resolved ~ victim_age + victim_sex + victim_race, data = .x, family = "binomial")),
    result = purrr::map(fit, broom::tidy)
  ) %>% 
  unnest(result) %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(OR_estimate = exp(estimate),
         OR_ci_lower = exp(estimate + qnorm(0.025) * std.error),
         OR_ci_upper = exp(estimate + qnorm(0.975)* std.error)) %>% 
  select(city_state, OR_estimate, OR_ci_lower, OR_ci_upper)  

fit_log_result %>% knitr::kable(digits = 3)
```

Create a plot that shows the estimated ORs and CIs for each city.

```{r}
fit_log_result %>% 
  mutate(CI = OR_ci_upper - OR_ci_lower,
         city_state = fct_reorder(city_state, OR_estimate)) %>% 
  ggplot(aes(x = city_state, y = OR_estimate)) +
  geom_point() +
  geom_errorbar(aes(x = city_state, ymin = OR_ci_lower, ymax = OR_ci_upper))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) +
   labs(
    title = "Logistic Estimated Odds Ratio for Solving Homicides (Male vs. Female) for Each City with 95% CI",
    x = "City, State",
    y = "Logistic Estimated Odds Ratio (Male vs. Female)")
```

Most of the cities have a logistic estimated odds ratio between 0 and 1 for solving homicides comparing male victims to female victims. `Fresno_CA`, `Stockton_CA` and `Albuquerque_NM` have larger predictions between 1 and 2. The 95 percent CI for these three cities are also larger than other cities indicating larger standard error. 

## Problem 3

Load and clean the data for regression analysis by converting numeric variables to appropriate factor variables. There is no missing data in `bwt_df`.
```{r}
bwt_df = read_csv("data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = factor(babysex),
    frace = factor(frace),
    malform = factor(malform),
    mrace = factor(mrace)
  ) 
```
Propose a regression model for birthweight using backward model selection. 

```{r}
fit0 = lm(bwt ~., data = bwt_df) 
step(fit0, direction = "backward") %>% broom::tidy()
```

The final linear regression model will include predictors of the following:
    `babysex`: baby’s sex (male = 1, female = 2),
    `bhead`: baby’s head circumference at birth (centimeters)
    `blength`: baby’s length at birth (centimeteres),
    `fincome`: family monthly income (in hundreds, rounded),
    `delwt`:  mother’s weight at delivery (pounds),
    `gaweeks`: gestational age in weeks 
    `mheight`: mother’s height (inches), 
    `mrace`: mother’s race ,
    `parity`:  number of live births prior to this pregnancy, 
    `ppwt`: mother’s pre-pregnancy weight (pounds), and 
    `smoken`:  average number of cigarettes smoked per day during pregnancy.
    
```{r}
fit1 = lm(bwt ~ babysex + bhead + blength + fincome + delwt + gaweeks + mheight + mrace + parity + ppwt + smoken, data = bwt_df)
```

Making a plot of model residuals against fitted values using `add_predictions` and `add_residuals`.

```{r}
bwt_df %>% 
  add_predictions(fit1) %>% 
  add_residuals(fit1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs. Fitted Values"
  )
```

Building another two models to predict the birthweight. 

```{r}
fit2 = lm(bwt ~ blength + gaweeks, data = bwt_df) 
fit3 = lm(bwt ~ bhead + blength + babysex + bhead * blength * gaweeks, data = bwt_df)
```

Comparing three models using cross validation. 

```{r}
cv_df =
  crossv_mc(bwt_df, 100) %>% 
  mutate(
    train = map(train, as_tibble), 
    test = map(test, as_tibble)
  ) %>% 
  mutate(
    fit1 = map(.x = train, ~lm(bwt ~ babysex + bhead + blength + fincome + delwt + gaweeks + mheight + mrace + parity + ppwt + smoken, data = .x)), 
    fit2 = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    fit3 = map(.x = train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength * gaweeks, data = .x))
  ) %>% 
  mutate(
    rmse_fit1 = map2_dbl(.x = fit1, .y = test, ~rmse(model = .x, data = .y)),
    rmse_fit2 = map2_dbl(.x = fit2, .y = test, ~rmse(model = .x, data = .y)),
    rmse_fit3 = map2_dbl(.x = fit3, .y = test, ~rmse(model = .x, data = .y))
  )
```

```{r}
cv_df %>% 
  summarize(
     rmse_fit1_avg = mean(rmse_fit1),
     rmse_fit2_avg = mean(rmse_fit2),
     rmse_fit3_avg = mean(rmse_fit3)
  )
```

The best model is `fit1` of linear regression with 11 predictors built by backward method. The second best model is `fit3` using head circumference, length, sex, and all interactions (including the three-way interaction) between these. The worst model is `fit2` using length at birth and gestational age as predictors (main effects only). 



